---
title: 'Tidytuesday: video games'
author: Philippe Massicotte
date: '2019-07-30'
slug: tidytuesday-video-games
categories:
  - R
  - Tidytuesday
tags: []
type: ''
subtitle: ''
image: ''
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  cache = FALSE,
  dpi = 300,
  out.width = "100%",
  fig.align = "center",
  fig.width = 8,
  fig.asp = 0.618, # 1 / phi
  fig.show = "hold",
  dev = "svg",
  message = FALSE
)

library(tidyverse)
library(ggpmthemes)
library(glue)
theme_set(theme_poppins())
```

I must admit, I played a lot of PC video games when I was younger, *Battlefield*, *Half-life*, *Dark Age of Camelot*, *World of Warcraft*, *Diablo* just to name a few. This is why this week [tidytuesday](https://github.com/rfordatascience/tidytuesday) was a good occasion to participate in this weekly R visualization challenge. 

## Video Games Dataset

> This week's data comes courtesy of Liza Wood via Steam Spy. She recently published a blog post on her data analysis of this video game data. She was kind enough to provide a fairly clean dataset, and I have done some small additional clean up seen below. There is time played, ownership, release date, publishing information, and for some a metascore! Lots of ways to slice and dice this data!

Let us get started! First, read the data and remove duplicated entries.

```{r}
video_games <- read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-07-30/video_games.csv") %>%
  mutate(release_date = as.Date(release_date, "%b %d, %Y")) %>%
  distinct(game, developer, publisher, .keep_all = TRUE)
```

## Top played games

Which games have the highest average playing time in the past two weeks? In the following graph, it can be seen that *Clicker Heroes* (released in 2015) has an average playtime of about 80 hours. That is 40 hours per week, a full-time job.

```{r}
video_games %>%
  top_n(16, average_playtime) %>%
  mutate(game = glue("{game} ({lubridate::year(release_date)})")) %>%
  mutate(game = fct_reorder(game, average_playtime)) %>%
  ggplot(aes(x = game, y = average_playtime / 60)) +
  geom_col() +
  coord_flip() +
  xlab(NULL) +
  ylab("Average played time (hours)") +
  labs(title = str_wrap("Average played time for the last two weeks", 25)) +
  labs(subtitle = "Only the top 16 averaged played game are shown")
```

## Temporal evolution of metascore

What is the evolution of the metascore by the publishers? It seems that *high ranked* publishers have pretty constant metascore for their games. However, there is an interesting decreasing trend with *SEGA* that started in 2006.

```{r}
equal_breaks <- function(n = 3, s = 0.05, ...) {
  function(x) {
    # rescaling
    d <- s * diff(range(x)) / (1 + 2 * s)
    seq(min(x) + d, max(x) - d, length = n)
  }
}

video_games %>%
  drop_na(metascore) %>%
  add_count(publisher) %>%
  filter(dense_rank(desc(n)) <= 6) %>%
  group_by(year = lubridate::year(release_date), publisher) %>%
  summarise(mean_metascore = mean(metascore), sd_metascore = sd(metascore)) %>%
  ggplot(aes(x = year, y = mean_metascore)) +
  geom_line(size = 2) +
  facet_wrap(~publisher, scale = "free_x") +
  scale_x_continuous(
    labels = function(x) floor(x),
    breaks = equal_breaks(n = 4, s = 0.05)
  ) +
  xlab(NULL) +
  ylab("Median metascore") +
  theme(legend.position = "none") +
  theme(panel.spacing = unit(2, "lines")) +
  labs(title = "Time series of metascore by publisher") +
  labs(subtitle = "Only the six publishers with the highest number of release are shown")

ggsave(
  here::here("graphs", "tidytuesday_2019_week31.png"),
  type = "cairo",
  dpi = 600,
  height = 8 * 0.6,
  width = 8
)
```

## Price evolution of games

The median prices of the released game appear to decrease between 2015 and 2018. Also, we can see the lowest prices are happening in January, right after Christmas.

```{r}
video_games %>%
  drop_na(release_date) %>%
  group_by(year = lubridate::year(release_date), month = lubridate::month(release_date, label = TRUE)) %>%
  summarise(medan_price = median(price, na.rm = TRUE)) %>%
  ggplot(aes(x = year, y = month, fill = medan_price)) +
  geom_tile() +
  scale_fill_viridis_c(option = "A", labels = scales::dollar) +
  coord_equal() +
  scale_x_continuous(expand = c(0, 0), breaks = seq(2000, 2020, by = 2)) +
  scale_y_discrete(expand = c(0, 0)) +
  labs(fill = "Median\nprice (USD)") +
  xlab(NULL) +
  ylab(NULL) +
  labs(title = "Price evolution of games") +
  labs(subtitle = "Median price calculated monthly and yearly")
```

There is a clear peak in November 2006 which can be explained by few game prices above 10$.

```{r}
video_games %>%
  filter(
    lubridate::year(release_date) == 2006 &
      lubridate::month(release_date) == 11
  ) %>%
  select(game, release_date, price) %>%
  arrange(desc(price)) %>%
  knitr::kable()
```

## Most expensive games

I was also surprised to see that the most expensive game was almost 600$ USD!

```{r}
video_games %>%
  top_n(5, price) %>%
  mutate(game = glue("{game} ({lubridate::year(release_date)})")) %>%
  mutate(game = str_wrap(game, 30)) %>%
  mutate(game = fct_reorder(game, price)) %>%
  ggplot(aes(x = game, y = price)) +
  geom_col() +
  coord_flip() +
  xlab(NULL) +
  ylab("Price (USD)") +
  labs(title = "Top priced games") +
  labs(subtitle = "Only shows the top 5 most expensive games") +
  scale_y_continuous(labels = scales::dollar)
```
