---
title: Tidytuesday (2019-08-06)
author: Philippe Massicotte
date: '2019-08-06'
slug: tidytuesday-2019-08-06
categories: []
tags:
  - Tidytuesday
  - Tidyverse
  - Bob Ross
type: ''
subtitle: 'Bob Ross - painting by the numbers'
description: 'test'
image: ''
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  cache = FALSE,
  dpi = 300,
  out.width = "100%",
  fig.align = "center",
  fig.width = 8,
  fig.asp = 0.618, # 1 / phi
  fig.show = "hold",
  dev = "svglite",
  message = FALSE,
  warning = FALSE,
  echo = TRUE
)

library(tidyverse)
library(ggpmthemes)
library(glue)
library(ggforce)

theme_set(theme_poppins())
```

This week [TidyTuesday](https://github.com/rfordatascience/tidytuesday/tree/master/data/2019/2019-08-06) is entitled *Bob Ross - painting by the numbers*. To cite [fivethirtyeight](https://fivethirtyeight.com/features/a-statistical-analysis-of-the-work-of-bob-ross/):

> Bob Ross was a consummate teacher. He guided fans along as he painted “happy trees,” “almighty mountains” and “fluffy clouds” over the course of his 11-year television career on his PBS show, “The Joy of Painting.” In total, Ross painted 381 works on the show, relying on a distinct set of elements, scenes and themes...

<div style="width:75%;height:0;padding-bottom:77%;position:relative;"><iframe src="https://giphy.com/embed/AbPNdmgk6TJK" width="100%" height="100%" style="position:absolute" frameBorder="0" class="giphy-embed" allowFullScreen></iframe></div><p><a href="https://giphy.com/gifs/bob-ross-AbPNdmgk6TJK">via GIPHY</a></p>

## The data 

The dataset consists of various keywords describing the content of Bob Ross paintings.

```{r}
bob_ross <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-08-06/bob-ross.csv") %>%
  janitor::clean_names() %>%
  mutate(title = str_to_title(title))

bob_ross <- bob_ross %>%
  select(-contains("frame")) %>%
  pivot_longer(
    -c(episode, title),
    names_to = "keyword",
    values_to = "is_present"
  ) %>%
  separate(episode, into = c("season", "episode"), sep = "E") %>%
  mutate(season = str_extract(season, "[:digit:]+")) %>%
  mutate_at(vars(season, episode), parse_number) %>%
  mutate(keyword = str_replace(keyword, "_", " ")) %>%
  mutate(keyword = str_to_title(keyword))

bob_ross
```

There are a total of `r n_distinct(bob_ross$keyword)` keywords used to describe his paintings. Upon my first exploration, I found some keywords were duplicated between words in singular and plural:

- lake/lakes
- mountain/mountains
- tree/trees

```{r}
bob_ross %>%
  distinct(keyword) %>%
  pull(keyword)
```
Let's do some more cleaning to the data.

```{r}
bob_ross <- bob_ross %>%
  mutate(keyword = case_when(
    keyword == "Lake" ~ "Lakes",
    keyword == "Mountain" ~ "Mountains",
    keyword == "Tree" ~ "Trees",
    TRUE ~ keyword
  ))

bob_ross <- bob_ross %>%
  group_by(season, episode, title, keyword) %>%
  summarise(is_present = sum(is_present)) %>%
  mutate(is_present = ifelse(is_present >= 1, 1, 0)) %>%
  ungroup()

bob_ross %>%
  distinct(keyword) %>%
  pull()
```

## The painting of Bob Ross

### Most painted objects

```{r, fig.height=7, fig.asp=NA}
top_painting_elements <- bob_ross %>%
  group_by(keyword) %>%
  summarise(n_painting = sum(is_present), total_painting = n_distinct(title)) %>%
  mutate(percent_object = n_painting / total_painting) %>%
  mutate(keyword = glue("{keyword} ({round(percent_object * 100, digits = 0)}%)")) %>%
  mutate(keyword = fct_reorder(keyword, percent_object))

top_painting_elements %>%
  filter(percent_object >= 0.1) %>%
  ggplot(aes(x = keyword, y = percent_object)) +
  geom_col() +
  coord_flip() +
  scale_y_continuous(labels = scales::percent, expand = expand_scale(mult = c(0, 0.1))) +
  xlab(NULL) +
  ylab("Percentage (%)") +
  labs(title = "Bob Ross most painted elements") +
  labs(subtitle = "Only showing items present more than 10% of the time") +
  theme(
    panel.grid = element_line(colour = NULL),
    panel.grid.minor = element_blank(),
    axis.ticks.y = element_blank(),
    axis.line = element_blank(),
    panel.border = element_blank()
  )
```


```{r}
top_painting_element <- bob_ross %>%
  group_by(keyword) %>%
  summarise(n = sum(is_present)) %>%
  ungroup() %>%
  filter(dense_rank(desc(n)) < 10) %>%
  inner_join(bob_ross) %>%
  group_by(season, keyword) %>%
  summarise(n = sum(is_present)) %>%
  ungroup() %>%
  arrange(desc(n))
```

### Trends in most painted elements

Are there any trends in the most painted elements?

```{r, fig.height = 7, fig.asp=NA, dev='jpeg'}

df <- top_painting_element %>%
  distinct(keyword) %>%
  arrange(keyword) %>%
  mutate(icon = case_when(
    keyword == "Bushes" ~ emoji("herb"),
    keyword == "Clouds" ~ emoji("cloud"),
    keyword == "Conifer" ~ emoji("palm_tree"),
    keyword == "Deciduous" ~ emoji("deciduous_tree"),
    keyword == "Grass" ~ emoji("seedling"),
    keyword == "Lakes" ~ emoji("ocean"),
    keyword == "Mountains" ~ emoji("mount_fuji"),
    keyword == "River" ~ emoji("ocean"),
    keyword == "Trees" ~ emoji("evergreen_tree")
  )) %>%
  inner_join(top_painting_element) %>%
  mutate(keyword = paste(keyword, icon))

df %>%
  mutate(keyword = fct_reorder(str_to_title(keyword), desc(n), .fun = sum)) %>%
  ggplot(aes(x = season, y = n)) +
  geom_line(size = 1) +
  facet_wrap(~keyword) +
  theme(legend.position = "none") +
  xlab("Season #") +
  ylab("Number of painted elements") +
  labs(title = "Most popular painted elements") +
  labs(subtitle = "Based on the nine most painted elements for all the episodes")
```

Heat map showing all the `r n_distinct(bob_ross$keyword)` elements in the database.

```{r, fig.height=8, fig.asp=NA}
bob_ross %>%
  group_by(season, keyword) %>%
  summarise(n = sum(is_present)) %>%
  ungroup() %>%
  mutate(keyword = str_replace(keyword, "_", " ")) %>%
  mutate(keyword = str_to_title(keyword)) %>%
  mutate(keyword = fct_reorder(keyword, n, sum)) %>%
  ggplot(aes(x = season, y = keyword, fill = n)) +
  geom_tile() +
  scale_fill_viridis_c(option = "plasma", labels = function(x) round(x)) +
  scale_x_continuous(expand = c(0, 0), breaks = seq(0, 35, by = 5)) +
  scale_y_discrete(expand = c(0, 0)) +
  xlab("Season #") +
  ylab(NULL) +
  labs(fill = "Number of painted element") +
  theme(legend.position = "bottom") +
  guides(fill = guide_colourbar(title.position = "top", title.hjust = 0.5)) +
  labs(title = "Heat map of all painted elements") +
  labs(subtitle = "Over all the season") +
  theme(
    panel.border = element_blank(),
    axis.ticks = element_blank()
  )

ggsave(
  here::here("graphs", "tidytuesday_2019_week32.png"),
  type = "cairo",
  dpi = 600,
  height = 8,
  width = 8
)
```

