---
title: Tidytuesday (2019-08-13)
author: Philippe Massicotte
date: '2019-08-14'
slug: tidytuesday-2019-08-13
categories: []
tags:
  - Tidytuesday
  - Tidyverse
type: ''
image: ''
subtitle: 'Roman Emperors'
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  cache = FALSE,
  dpi = 300,
  out.width = "100%",
  fig.align = "center",
  fig.width = 8,
  fig.asp = 0.618, # 1 / phi
  fig.show = "hold",
  dev = "svglite",
  message = FALSE,
  warning = FALSE,
  echo = TRUE
)

library(tidyverse)
library(ggpmthemes)
library(glue)
library(ggforce)

theme_set(theme_poppins())
```

This week, the [TidyTuesday challenge](https://github.com/rfordatascience/tidytuesday/tree/master/data/2019/2019-08-13) is about Roman emperors. The dataset provides the name, birth, death and reign information for 68 Roman emperors.

## The data

```{r}
emperors <-
  readr::read_csv(
    "https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-08-13/emperors.csv"
  )

emperors
```

## Exploratory graph

In the following graph, I wanted to visualize the relationship between the age and the reign duration of each emperor. First, lets calculate the age and the reign period of each emperor.

```{r}
emperors <- emperors %>%
  drop_na(death, birth, reign_start, reign_end) %>%
  filter(death >= birth) %>%
  mutate(age = lubridate::time_length(death - birth, unit = "year")) %>% 
  mutate(reign_duration = lubridate::time_length(reign_end - reign_start, unit = "year")) 
```

I can now graph the relationship between the age and the reign duration of each emperor.

```{r, fig.height=14, fig.asp=NA}
emperors %>% 
  mutate(
    name = glue(
      "{name} ({round(age, digits = 0)} years old)"
    )
  ) %>%
  mutate(name = fct_reorder(name, death)) %>%
  ggplot(aes(
    x = birth,
    xend = death,
    y = name,
    yend = name
  )) +
  geom_segment(size = 3, lineend = "round", aes(color = dynasty)) +
  geom_segment(
    aes(x = reign_start, xend = reign_end),
    color = "black",
    size = 1,
    lineend = "round"
  ) +
  geom_label(aes(
    x = death,
    y = name,
    label = glue(
      "Reign: {round(reign_duration, digits = 2)} years"
    )
  ),
  hjust = -0.1,
  label.size = 0,
  size = 3,
  family = "Poppins"
  ) +
  geom_mark_circle(
    aes(
      label = glue("{name}"),
      description = "The oldest emperor",
      filter = age == max(age),
      color = dynasty
    ),
    expand = unit(0.01, "mm"),
    label.family = c("Poppins"),
    label.fontsize = 8,
    label.buffer = unit(5, "mm"),
    label.hjust = 1, 
    show.legend = FALSE
  ) +
  geom_mark_circle(
    aes(
      label = glue("{name}"),
      description = "The longest reign",
      filter = reign_duration == max(reign_duration),
      color = dynasty
    ),
    expand = unit(0.01, "mm"),
    label.family = c("Poppins"),
    label.fontsize = 8,
    label.buffer = unit(5, "mm"),
    label.hjust = 1,
    show.legend = FALSE
  ) +
  theme(legend.position = "bottom") +
  theme(panel.grid.major.x = element_blank(), panel.grid.minor.x = element_blank()) +
  scale_x_date(expand = expand_scale(mult = c(0.02, 0.3))) +
  ylab(NULL) +
  xlab("Years") +
  labs(color = "Dynasty") +
  labs(title = str_wrap("Roman emperors: their birth, death and reign", 35)) +
  labs(subtitle = str_wrap("Colored bars represent the lifespan whereas the black bars the duration of the reign of each emperor", 60)) +
  labs(caption = "Data: https://bit.ly/2N2spWr | Visualization: @philmassicotte") +
  scale_color_brewer(palette = "Pastel2")

ggsave(
  here::here("graphs", "tidytuesday_2019_week33.png"),
  type = "cairo",
  dpi = 600,
  height = 14,
  width = 8
)
```

## Causes of death

```{r}
emperors %>% 
  count(cause) %>% 
  mutate(cause = fct_reorder(cause, n)) %>% 
  ggplot(aes(x = cause, y = n)) +
  geom_col() +
  coord_flip() +
  xlab(NULL) +
  ylab("Number of death") +
  scale_y_continuous(expand = expand_scale(mult = c(0, 0.2))) +
  labs(title = "Principal causes of death") +
  labs(subtitle = glue("Based on {nrow(emperors)} observations")) +
  labs(caption = "Data: https://bit.ly/2N2spWr | Visualization: @philmassicotte")
```

## Some stats

```{r, echo=FALSE}
emperors %>%
  summarise(
    `Average age at death` = paste(round(mean(age), 0), "years old"),
    `Average reign duration` = paste(round(mean(reign_duration), 0), "years old"),
    `Youngest emperor` = paste(round(min(age), 0), "years old"),
    `Oldest emperor` = paste(round(max(age), 0), "years old"),
    `Longest reign` = paste(round(max(reign_duration), 2), "years"),
    `Shortest reign` = paste(round(min(reign_duration), 2), "years"),
  ) %>%
  gather() %>% 
  set_names(c("Stats", "Age")) %>% 
  knitr::kable()
```

